Error Reporting Parsers: a Monad Transformer Approach

### 2013-06-10 article.tex
Add monoids
Signed-off-by: Edward Z. Yang <ezyang@mit.edu>
example code showing that ErrorT/MaybeT transformer order doesn't matter

@@

On the other hand, the order of Error and Maybe relative to each other isn't 
important.  But don't just take my word for it -- try it for yourself!
\EZY{It's worth mentioning why, formally speaking, this is the case, which has
to do with the relevant structures being isomorphic to each other.}
\Matt{Uh-oh!  I don't really know anything formal about monad transformers,
I really have no idea how to go about showing something like this!
I guess I'll have to rethink my goals for this section.}
\EZY{It's pretty simple, and just has to do with the fact that Maybe (Either a b) is
equivalent (modulo laziness) with Either a (Maybe b).  To show this is a case, you have
to define functions which convert from one to the other, and then show that the roundtrip
is identity.}

Another important point to note is that our parsers don't specify
which order -- Maybe/State or State/Maybe -- they prefer, and will 
work with either.  This means that the semantics are ultimately determined by 
the actual monad stack we use.

>>>>

\newcommand{\Matt}[1]{\Red{Matt: #1}}

@@

On the other hand, the order of Error and Maybe relative to each other isn't 
important.  We can show this by losslessly converting values from one stack to 
the other and back:
\begin{verbatim}
type Stack1 e a = Either e (Maybe a)
type Stack2 e a = Maybe (Either e a)

forward :: Stack1 e a -> Stack2 e a
forward (Left e)          =  Just $ Left e
forward (Right Nothing)   =  Nothing
forward (Right (Just x))  =  Just $ Right x

backward :: Stack2 e a -> Stack1 e a
backward Nothing           =  Right Nothing
backward (Just (Left e))   =  Left e
backward (Just (Right x))  =  Right $ Just x

identityForward :: Stack1 e a -> Stack1 e a
identityForward = backward . forward

identityBackward :: Stack2 e a -> Stack2 e a
identityBackward = forward . backward
\end{verbatim}

No partial functions were used.  Now we can use these functions to
convert between the two stacks.  Note the final result is the same as the input:
\begin{verbatim}
vals = [Left "an error", Right Nothing, Right $ Just "success"]

*Main> map forward vals
[Just (Left "an error"),Nothing,Just (Right "success")]

*Main> map identityForward vals
[Left "an error",Right Nothing,Right (Just "success")]
\end{verbatim}

Another important point is that our parsers don't specify
which order -- Maybe/State or State/Maybe -- they prefer, and will 
work with either.  This means that the semantics are ultimately determined by 
the actual monad stack we use.
